### 2019-09-25 slides.tex
Update on Overleaf.

\input{header_slides.tex}

\begin{document}

\title[Agent-based modeling city networks]{An agent-based model for city networks based on interactions between firms}
\author[Raimbault \and Zdanowska]{J.~Raimbault$^{1,2,3\ast}$ \and N. Zdanowska$^{1,3}$\\\medskip
$^{\ast}$\texttt{j.raimbault@ucl.ac.uk}
}

\institute[UCL]{$^{1}$CASA, UCL\\
$^{2}$UPS CNRS 3611 Complex Systems Institute Paris\\
$^{3}$UMR CNRS 8504 G{\'e}ographie-cit{\'e}s
}

\date[2nd October 2019]{CCS 2019\\
Satellite Simulation models\\
October 2nd 2019
}

\frame{\maketitle}

\section{Introduction}

\sframe{Urban firm networks}{

General framework:

\begin{itemize}
    \item Cities as maximisation of socio-economic interactions 
    \item Network economies \cite{sassen1991global} \cite{castells1996information}
    \item World city network driven by interactions between firms (Taylor, 2001; Martinus and Sigler, 2018)
    \item Asymmetrical spatial interactions -- different relative position of cities
\end{itemize}

\bigskip

}

\sframe{An evolutionary approach to Urban Systems}{

\begin{itemize}
    \item Evolutionary Theory of Urban Systems \cite{pumain1997pour} \cite{pumain2006villes}
    \item Adaptive cycles and diffusion on innovation (H{\"a}gerstrand, 1973) 
    \item Path dependence \cite{martinsunley2006information}
    \item Selection and emerging structures of systems
\end{itemize}

\bigskip
% discussion theorie Denise
}

\sframe{Firm linkages structuring urban systems}{

% examples de resultats pour l'Europe Centrale et Orientale

 \begin{center}
 \includegraphics[width=0.7\textwidth]{figures/Figure1.png}
 \end{center}

  \textit{Central and Eastern European cities within ownership links of firms in 2013.}

}

% si besoin je peux rajouter d'autres exemples sur le UK ou carrement des trucs de Celine sur les reseaux de firmes

\sframe{Urban Firm linkages and geo-economic processes}{

Assumptions

\begin{itemize}
    \item Metropolisation vs. regionalisation effects
    \item Relationship local/global of multinational firm linkages 
    \item Specialisation-driven factors and drivers of innovation
    \item Macroeconomic exogenous chocs and resilience of urban systems 
\end{itemize}

How can we approach geographical and economic processes within urban networks of firms with a generative model?

}
\sframe{Towards a generative model of urban firm networks}{

Towards a generative model of urban firm networks

% research question / objective
%

}

\section{Model}

\sframe{Model rationale}{

}

\sframe{Formalization}{

Cities characterized by economic size $E_i$ (GDP) and economic structure $s_{ik}$

\bigskip

Starting from an initial network, at each time step:
\begin{enumerate}
    \item Evolve cities with an interaction model (\cite{raimbault2018indirect} or \cite{cottineau2015growing})
    \item Add a fixed number of links randomly, following a probability function of sizes, sector proximity, and geographical and socio-cultural proximities
\end{enumerate}

}

\sframe{Formalization}{

Probability for a new link follows a generalized Cobb-Douglas function

\begin{equation}
     p_{ij} \propto \left(\frac{E_{i}}{E}\right)^{\gamma_F} \cdot \left(\frac{E_{j}}{E}\right)^{\gamma_T} \cdot \left(\frac{w_{ij}}{W}\right)^{\gamma_W} \cdot s(S_{ik},S_{jk})^{\gamma_S} \cdot \exp \left(- \gamma_D \cdot d_{ij}\right) \cdot \exp \left(- \gamma_G \cdot g_{ij}\right)
\end{equation}

}

\sframe{Model indicators}{

Geographical indicators:

\begin{itemize}
    \item Internationalisation
    \item Metropolisation
    \item Regionalisation
    \item Specialisation
\end{itemize}

\bigskip

Network and flows indicators

}

\sframe{Simulation on synthetic systems of cities}{

% la ou avant il faut surement expliquer que c'est pas totalement synthetique, que tu pars de X villes europeennes de plus de 50 000 habitants (pour les plus nuls comme moi)
% si c'est completement synthetiques, l'histoire des villes europeenes c'est pour avoir les ordres de grandeur de la hierarchie et du nombre de villes, mais c'est pas parametrise sur des vraies donnees - mais oui ca ca sera explique dans comment construire le systeme de ville synthetique.

% ok desolee je n'avais pas bien compris

}

\section{Results}

\sframe{Implementation}{

% for now no dynamics of cities

}

\sframe{Simulation of urban networks}{

    % examples

    \centering

    \includegraphics[width=0.48\textwidth]{figures/ex_alleq-highgravity_seed-12102_t1500.png}\hspace{0.1cm}
    \includegraphics[width=0.48\textwidth]{figures/ex_alleq-lowgravity_seed-12102_t1500.png}

}

\sframe{Statistical convergence}{

Good convergence for most indicators: average sharpes ratios on a one-factor sampling above 5 (except for metropolization index $\simeq 1.5$)

% + histograms

}

\sframe{Interaction decay}{

    % One factor sampling results
    %  20190924_162740_ONEFACTOR_REPLICATIONS_SYNTHETIC_GRID

    \includegraphics[width=0.48\textwidth]{figures/internationalization-gravityDecay_errorbars.png}
    \includegraphics[width=0.48\textwidth]{figures/rhoDegreeSize-gravityDecay_errorbars.png}
}

\sframe{Sector proximity}{

    % One factor sampling results
    %  20190924_162740_ONEFACTOR_REPLICATIONS_SYNTHETIC_GRID

     \includegraphics[width=0.48\textwidth]{figures/internationalization-gammaSectors_errorbars.png}
    \includegraphics[width=0.48\textwidth]{figures/rhoDegreeSize-gammaSectors_errorbars.png}

}

\sframe{Grid exploration}{

}

% other experiments ? PSE on correlations ?

\section{Discussion}

% - future work etc (classical discussion)
% - on the role of model exploration and simulation to gain knowledge from the model

% further comparing with real data?

\sframe{On the role of model exploration}{

}

\sframe{Discussion}{

}

\sframe{Conclusion}{

$\rightarrow$

\bigskip
\bigskip

\textbf{Open repository} at

\texttt{https://github.com/JusteRaimbault/ABMCitiesFirms}

\bigskip

\textbf{Acknowledgments}: thanks to the \textit{European Grid Infrastructure} for access to the infrastructure.

}

\sframe{Reserve slides}{

\centering

\Large

\textbf{Reserve Slides}

}

%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[allowframebreaks]
\frametitle{References}
\bibliographystyle{apalike}
\bibliography{biblio}
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\end{document}