### 2020-03-04 readme_amadeus.tex

\documentclass[10pt]{article}
\usepackage[margin=2cm]{geometry}

\begin{document}

\title{Amadeus Van Dijk database}

\maketitle

\section*{Synthesis}

\subsection*{Table nodes}

\subsection*{Table links}

\subsection*{Counts}

\begin{itemize}
    \item 3,053,540 firms
    \item 2,715,188 with coordinates
    \item 2,653,434 within Europe, 2,653,034 within EU (due to the way Amadeus is constructed)
    \item 1,920,909 within FUAs
    \item 1,495,651 within FUAs with a link
    \item 753,224 within FUAs, linked and with turnover
\end{itemize}

\begin{itemize}
    \item 1,866,936 links
    \item 794,782 with an origin European country, 1,691,007 with a destination European country, 761,173 with both % most links are intercontinental ! we get with the amadeus either links from EU, or to EU -> international study interesting ? 
    \item 616,031 with an origin FUA, 1,254,374 with a destination FUAs
    \item 555,196 between FUAs % FIXME proper join
    \item 161,303 between FUAs, with ownership \% information and destination turnover
    \item when aggregating at FUA (without years), 4,978 links (on 451,584 possible, very sparse)
\end{itemize}

\subsection*{FUA Network}

Quite modular (0.35); fixed effect by country significant

\end{document}