### 2020-09-01 S1_Text.tex

\input{si.tex}
\begin{document}
\vspace*{0.35in}
\justify

\section*{S1 Text: UK's economic sectors dependent on international links}

Based on \emph {AMADEUS} dataset, the most vulnerable economic sectors in the United Kingdom have been highlighted. The later are the ones due to be mostly affected by eventual future lock-downs of the international economy (international ownership links).

When considering the UK sectors generating the highest turnover in 2018 mostly dependent on foreign capital they own (out-going links from the UK) these are extraction of crude petroleum, manufacture of tobacco products and of basic pharmaceutical goods. In terms of in-going links to the UK, activities of head offices and holding companies, sale of cars and manufacture of motor vehicles are sectors mostly dependent on foreign ownership coming from abroad.

\FloatBarrier

\begin{table}[!ht]
\begin{adjustwidth}{-2.25in}{0in}
\caption{Top 10 UK industries depending on foreign ownership expressed in total turnover (out-going links from the UK)\label{tab:uk_outgoing}}
\centering
\medskip
\begin{tabular}{|l|c|c|c|c|}
\hline
Classification  & NACE Industries & Total turnover \\ 
\hline
1 &     Extraction of crude petroleum &   96,458,258,642  \\
\hline
2 & Manufacture of tobacco products &  5,477,272,464  \\
\hline
3 & Manufacture of basic pharmaceutical products &  4,825,881,930 \\
\hline
4 &  Retail sale in non-specialised stores with food, beverages or tobacco predominating  &   3,681,373,374  \\
\hline
5 &  Wireless telecommunications activities &   3,379,297,218  \\
\hline
6 & Event catering activities & 2,871,593,754  \\
\hline
7 &  Other mining and quarrying n.e.c. &  1,590,548,873  \\
\hline
8 & Business and other management consultancy activities &  1,394,373,014 \\
\hline
9 &  Activities of head offices &  1,322,692,365 \\
\hline
10 & Manufacture of air and spacecraft and related machinery &  1,292,065,497 \\
\hline
\end{tabular}
\end{adjustwidth}
\end{table}

\begin{table}[!ht]
\begin{adjustwidth}{-2.25in}{0in}
\caption{Top 10 UK industries depending on foreign ownership expressed in total turnover (in-going links to the UK)\label{tab:uk_ingoing}}
\centering
\medskip
\begin{tabular}{|l|c|c|c|c|}
\hline
Classification  & NACE Industries & Total turnover \\ 
\hline
1 &      Activities of head offices &  358,099,870  \\
\hline
2 & Activities of holding companies  & 259,008,123  \\
\hline
3 & Other business support service activities n.e.c. & 183,098,170 \\
\hline
4 &  Security and commodity contracts brokerage  &  72,482,167 \\
\hline
5 &  Manufacture of motor vehicles  &   65,162,757 \\
\hline
6 &  Sale of cars and light motor vehicles  &    61,019,026  \\
\hline
7 &  Other financial service activities, except insurance and pension funding & 47,003,937 \\
\hline
8 &  Other personal service activities n.e.c.  & 37,926,668  \\
\hline
9 &  Wholesale of solid, liquid and gaseous fuels and related products & 36,693,362 \\
\hline
10 &  Other information technology and computer service activities  &  35,668,779 \\
\hline
\end{tabular}
\end{adjustwidth}
\end{table}

\newpage
When expressed in total number of links, activities mostly dependent on out-going links from the UK are activities of holding companies, of head offices and fund management activities. In terms of in-going links these are business and consultancy activities and renting and operating of own and leased real estate.

\begin{table}[!ht]
\begin{adjustwidth}{-2.25in}{0in}
\caption{Top 10 UK industries depending on foreign ownership expressed in total links (out-going links from the UK)\label{tab:uk_outgoinglinks}}
\centering
\medskip
\begin{tabular}{|l|c|c|c|c|}
\hline
Classification  & NACE Industries & Total links \\ 
\hline
1 & Activities of holding companies &   686  \\
\hline
2 & Other business support service activities n.e.c.  &  567  \\
\hline
3 & Activities of head offices &  503 \\
\hline
4 &  Fund management activities &   407  \\
\hline
5 &  Business and other management consultancy activities &  398  \\
\hline
6 &  Extraction of crude petroleum &  366  \\
\hline
7 &  Other activities auxiliary to financial services, except insurance and pension funding & 240 \\
\hline
8 &  Development of building projects  & 221 \\
\hline
9 &  Manufacture of basic pharmaceutical products &  212 \\
\hline
10 &  Other information technology and computer service activities  &  181 \\
\hline
\end{tabular}
\end{adjustwidth}
\end{table}

\begin{table}[!ht]
\begin{adjustwidth}{-2.25in}{0in}
\caption{Top 10 UK industries depending on foreign ownership expressed in total links (in-going links to the UK)\label{tab:uk_ingoinglinks}}
\centering
\medskip
\begin{tabular}{|l|c|c|c|c|}
\hline
Classification  & NACE Industries & Total links \\ 
\hline
1 &      Other business support service activities n.e.c. &   16,140  \\
\hline
2 & Renting and operating of own or leased real estate  &  6,320  \\
\hline
3 & Business and other management consultancy activities &  6,094  \\
\hline
4 &  Activities of head offices &   5,628  \\
\hline
5 &  Computer consultancy activities &   5,047  \\
\hline
6 &  Activities of holding companies &  4,728  \\
\hline
7 &  Other personal service activities n.e.c. &  3,996 \\
\hline
8 &  Development of building projects &  3,946 \\
\hline
9 &  Other information technology and computer service activities &  3,757 \\
\hline
10 &  Other financial service activities, except insurance and pension funding  &  3,321 \\
\hline
\end{tabular}
\end{adjustwidth}
\end{table}

\end{document}