### 2020-09-11 S1_Text.tex
SuppMat

\input{si.tex}

\renewcommand{\thetable}{S1-\Alph{table}}

\begin{document}
\vspace*{0.35in}
\justify

\section*{S1 Text: UK's economic sectors dependent on international links}

Based on \emph {AMADEUS} dataset, the most vulnerable economic sectors in the United Kingdom have been highlighted. The later are the ones due to be mostly affected by eventual future lock-downs of the international economy (international ownership links).

When considering the UK sectors generating the highest turnover in 2018 mostly dependent on foreign capital they own (out-going links from the UK - see below Table~\ref{tab:uk_outgoing}) these are extraction of crude petroleum, manufacture of tobacco products and of basic pharmaceutical goods. In terms of in-going links to the UK (see below Table~\ref{tab:uk_ingoing}), activities of head offices and holding companies, sale of cars and manufacture of motor vehicles are sectors mostly dependent on foreign ownership coming from abroad.

\FloatBarrier

\begin{table}[!ht]
\begin{adjustwidth}{-2.25in}{0in}
\caption{Top 10 UK industries depending on foreign ownership expressed in total turnover (out-going links from the UK)\label{tab:uk_outgoing}}
\centering
\medskip
\begin{tabular}{|l|c|c|c|c|}
\hline
Classification  & NACE Industries & Total turnover \\ 
\hline
1 &     Extraction of crude petroleum &   96,458,258,642  \\
\hline
2 & Manufacture of tobacco products &  5,477,272,464  \\
\hline
3 & Manufacture of basic pharmaceutical products &  4,825,881,930 \\
\hline
4 &  Retail sale in non-specialised stores with food, beverages or tobacco predominating  &   3,681,373,374  \\
\hline
5 &  Wireless telecommunications activities &   3,379,297,218  \\
\hline
6 & Event catering activities & 2,871,593,754  \\
\hline
7 &  Other mining and quarrying n.e.c. &  1,590,548,873  \\
\hline
8 & Business and other management consultancy activities &  1,394,373,014 \\
\hline
9 &  Activities of head offices &  1,322,692,365 \\
\hline
10 & Manufacture of air and spacecraft and related machinery &  1,292,065,497 \\
\hline
\end{tabular}
\end{adjustwidth}
\end{table}

\begin{table}[!ht]
\begin{adjustwidth}{-2.25in}{0in}
\caption{Top 10 UK industries depending on foreign ownership expressed in total turnover (in-going links to the UK)\label{tab:uk_ingoing}}
\centering
\medskip
\begin{tabular}{|l|c|c|c|c|}
\hline
Classification  & NACE Industries & Total turnover \\ 
\hline
1 &      Activities of head offices &  358,099,870  \\
\hline
2 & Activities of holding companies  & 259,008,123  \\
\hline
3 & Other business support service activities n.e.c. & 183,098,170 \\
\hline
4 &  Security and commodity contracts brokerage  &  72,482,167 \\
\hline
5 &  Manufacture of motor vehicles  &   65,162,757 \\
\hline
6 &  Sale of cars and light motor vehicles  &    61,019,026  \\
\hline
7 &  Other financial service activities, except insurance and pension funding & 47,003,937 \\
\hline
8 &  Other personal service activities n.e.c.  & 37,926,668  \\
\hline
9 &  Wholesale of solid, liquid and gaseous fuels and related products & 36,693,362 \\
\hline
10 &  Other information technology and computer service activities  &  35,668,779 \\
\hline
\end{tabular}
\end{adjustwidth}
\end{table}

\newpage
When expressed in total number of links (Table~\ref{tab:uk_outgoinglinks} and Table~\ref{tab:uk_ingoinglinks}), activities mostly dependent on out-going links from the UK are activities of holding companies, of head offices and fund management activities. In terms of in-going links these are business and consultancy activities and renting and operating of own and leased real estate.

\begin{table}[!ht]
\begin{adjustwidth}{-2.25in}{0in}
\caption{Top 10 UK industries depending on foreign ownership expressed in total links (out-going links from the UK)\label{tab:uk_outgoinglinks}}
\centering
\medskip
\begin{tabular}{|l|c|c|c|c|}
\hline
Classification  & NACE Industries & Total links \\ 
\hline
1 & Activities of holding companies &   686  \\
\hline
2 & Other business support service activities n.e.c.  &  567  \\
\hline
3 & Activities of head offices &  503 \\
\hline
4 &  Fund management activities &   407  \\
\hline
5 &  Business and other management consultancy activities &  398  \\
\hline
6 &  Extraction of crude petroleum &  366  \\
\hline
7 &  Other activities auxiliary to financial services, except insurance and pension funding & 240 \\
\hline
8 &  Development of building projects  & 221 \\
\hline
9 &  Manufacture of basic pharmaceutical products &  212 \\
\hline
10 &  Other information technology and computer service activities  &  181 \\
\hline
\end{tabular}
\end{adjustwidth}
\end{table}

\begin{table}[!ht]
\begin{adjustwidth}{-2.25in}{0in}
\caption{Top 10 UK industries depending on foreign ownership expressed in total links (in-going links to the UK)\label{tab:uk_ingoinglinks}}
\centering
\medskip
\begin{tabular}{|l|c|c|c|c|}
\hline
Classification  & NACE Industries & Total links \\ 
\hline
1 &      Other business support service activities n.e.c. &   16,140  \\
\hline
2 & Renting and operating of own or leased real estate  &  6,320  \\
\hline
3 & Business and other management consultancy activities &  6,094  \\
\hline
4 &  Activities of head offices &   5,628  \\
\hline
5 &  Computer consultancy activities &   5,047  \\
\hline
6 &  Activities of holding companies &  4,728  \\
\hline
7 &  Other personal service activities n.e.c. &  3,996 \\
\hline
8 &  Development of building projects &  3,946 \\
\hline
9 &  Other information technology and computer service activities &  3,757 \\
\hline
10 &  Other financial service activities, except insurance and pension funding  &  3,321 \\
\hline
\end{tabular}
\end{adjustwidth}
\end{table}

\end{document}