### 2022-05-12 CoverLetter.tex
Update on Overleaf.

\documentclass[11pt,a4paper,sans]{moderncv}

\usepackage[document]{ragged2e}

\moderncvstyle{banking}   
\moderncvcolor{red} 
\renewcommand{\familydefault}{\rmdefault}
\usepackage[utf8]{inputenc}
\usepackage[scale=0.75]{geometry}

\firstname{}
\lastname{}
\begin{document}

\recipient{Editors PLoS ONE}{}
\date{May 12th 2022}
\opening{Dear Editors,}
\closing{Yours faithfully,\\
Juste Raimbault\\
LASTIG, Univ Gustave Eiffel, IGN-ENSG
}

\makelettertitle

\justify
We are pleased, with my co-authors Natalia Zdanowska (Luxembourg Institute of Socio-economic Research) and Elsa Arcaute (University College London), to submit an original research article entitled ``Modeling growth of urban firm networks'' for consideration for publication in PLoS ONE. This article introduces a generative model for urban firm networks, to understand underlying processes, with possible applications to public policies.

A previous version of this manuscript was submitted to PLoS ONE in September 2020, with the submission identifier PONE-S-20-35765. Due to diverse unforeseen circumstances including job uncertainty and instability through the pandemic, we were not able to revise the manuscript in time after receiving a first round of reviews suggesting a revision. This current submission thus corresponds to the revision of this previous submission, and we included the response to reviewers and manuscript with track changes in submitted files.

This manuscript has not been published and is not under consideration for publication elsewhere, nor contains previously published work. A preprint of a previous version was deposited on arxiv with the identifier 2009.05528.

We have no conflicts of interest to disclose. We do not have any opposed reviewer.

This paper was initially submitted to the PloS ONE collection ``Cities as Complex Systems''. We therefore suggest as editor D. Rybski (ORCID 0000-0001-6125-7705, Potsdam Institute for Climate Impact Research) who is on the editorial board and was editor of the collection. We also suggest as editor F. Biljecki (ORCID 000-0002-6229-7749, National University of Singapore).

Thank you for your consideration.
\justify

\makeletterclosing

\end{document}